### 2010-02-22 paper/semsearch10/paper.tex
latest

\documentclass{www2010-submission}
\begin{document}

\title{Leveraging Semantics for a Deep Web Question and Answering System} 
\numberofauthors{7}
\author{
	\alignauthor Christan Grant\\
	\email{cgrant@cise.ufl.edu}

	\alignauthor Clint P. George\\
	\email{clintpg@gmail.com}

	\alignauthor Joseph N. Wilson\\
	\email{jnw@cise.ufl.edu}

	\alignauthor Peter J. Dobbins\\
	\email{pjd@cise.ufl.edu}

	\alignauthor Jeff T. Depree\\
	\email{jepree@cise.ufl.edu}

	\alignauthor Christopher A. Shields\\
	\email{cshields@cise.ufl.edu}

	\alignauthor Joir-dan Gumbs\\
	\email{jgumbs@cise.ufl.edu}

	\sharedaffiliation
	\affaddr{Department of Computer and Information Science and Engineering }  \\
	\affaddr{University of Florida} \\
	\affaddr{Gainesville, Florida 32611-6120}
}
\date{April 26,2010}

\maketitle

\begin{abstract}
- It is difficult to answer question that change over time
- However, if the method of answering the question can be realized, the question can be easily answered
- We are building a system that supports a learning how to answer particular questions
- We can answer questions that are semantically similar to our store of questions
- We do this by building a semantic library ...
\end{abstract}

\category{H.4.m}{Information Systems Applications}{Miscellaneous}

\terms{SSQ}

\keywords{Questions Answering, re-finding}

\section{Introduction}
% Christan
- The best way to answer questions is to see/know if a question has been asked before.
- IEEE potentials talks about the the beaten trail.
- Use reuse in the specific domain of question answering

\section{Statistical Hierarchy}
\subsection*{Class Hierarchy}
%clint
\subsection*{Markov Blanket}
%joir-dan

\section{Semantic Query Representation}
% Christan - fois

\section{Building Query Store}
\subsection{Query Resolution Recorder}
% Chris
\subsection{Generalization}
% Christan

\section{Semantic Query Answering}
\subsection{Semi-Structured Query}
% Joir-dan
\subsection{SSQ Ranking}
% Clint
\subsection{Query Re-Execution}
% Chris

\section{Results}
% Walk through of example query

\section{Future Work}

\section{Conclusion}

\section{Acknowledgements}

\section{References}

\end{document}            % End of document.