### 2010-03-03 trunk/paper/semsearch10/content/results_futurework_conclusion.tex

\section{Results}
% Walk through of example query

\section{Future Work}
- Talk about the offline execution mode and class structural matching algorithm
- Researchers have stated 39, 57, and 63 percent of web queries are unique \cite{1277770,331405,621942}.  We believe that an even higher number of those are parameterizable, meaning, a set of queries are all related and may be answered using a single QRM.  We plan to investigate this.

% Sort classes based on the prior

\section{Conclusion}

\section{Acknowledgements}