### 2010-03-03 trunk/paper/semsearch10/paper_old_old.tex

\documentclass{www2010-submission}
\begin{document}

\title{Re-finding to support a Deep Web Question Answering System} 
\numberofauthors{1}
\author{
	 \alignauthor Christan Grant, Clint P. George, Joseph N. Wilson, Peter J. Dobbins, Jeff T. Depree, Christopher A. Shields and Joir-dan Gumbs\\
	 \affaddr{Dept. of Computer Science, University of Florida} \\ \affaddr{ Gainesville, Florida, USA} \\
\email{\{cgrant, cgeorge, jnw, pjd, jepree, cshields, jgumbs\} @cise.ufl.edu}
}
\date{April 26,2010}

\maketitle

\category{H.4.m}{Information Systems Applications}{Miscellaneous}

\terms{SSQ}

\keywords{Questions Answering, re-finding}

\section{Semantic Query Representation}
% Christan - fois
- We translate natural language queries into a format we call - semi structured query
- the semi structure query is formated as followed...
- this format especially allows us to better match deep web parameters with queries

\section{Establishing the Semantic Hierarchy}

\section{Building Query Store}

\section{Natural Language Queries}
\subsection{Semi-Structured Query}
% Joir-dan

%-------------------------------------------------------------------------------

\subsection{Query Re-Execution}
% Chris

%
% The bibliography for the citations in the paper 
\bibliographystyle{abbrv}
\bibliography{citations}  						% citations.bib is the name of the Bibliography

\end{document}            						% End of document.