### 2010-03-04 trunk/paper/semsearch10/content/introduction.tex

\section{Introduction}
% Christan
When traveling though a jungle to a destination, it is easy to get
lost.  The first person to take blaze a trail may make a number of
mistakes in trying to find the best path to any destination.  However,
those who come later can easily follow a well-marked trail without the
level of effort required by the path-finder. Olsen and Malizia
describe this idea as \emph{exploiting trails} \cite{5379671}.  The
Morpheus Question Answering System uses this insight to improve the
efficency of retrieval in semantic search.

Rather than treating each user's search as a unique entity and working
from first principles, we exploit the fact that a similar search may
have been performed already.
Researchers have shown that almost 40 percent of the all
queries are repetitions of previous queries.\cite{1277770}
This points to an  area for optimization for search engines.
We have built a prototype system for answering questions using deep-web sources
that employs previously answered questions to obtain new answers.