### 2010-03-04 paper/semsearch10/content/example.tex
third round of changes by jnw

%jnw

The end-user employs the Morpheus question-answering system much like a
search engine with a natural-language interface. One can type a question
in a text box and receive a response from the system. Those questions
dealing with an ontological realm familiar to the system, whose form is
similar to previously answered queries in that realm, and that can be
answered using (possibly deep web) information from a source the system
has encountered can yield relevant responses.

\begin{figure}
\caption{Typical Morpheus interaction}
\label{fig:screenshot}
\end{figure}

Figure \ref{fig:screenshot} shows a screenshot of a typical Morpheus
interaction. The user has typed a question ``Blah blah blah!!!!!!?'' and
received a response together with information describing assumptions
about the question, the likely relevance of the answer provided,
the source of information, and the method of answer production.

Each answer to a question is created using a \emph{query response
  method} (QRM). A QRM is constructed when a special user, known as a
\emph{guide} answers a question using an instrumented web browser (the
\emph{query response recorder} or QRR) that records web
interactions. The system extracts the relevant artifacts of the
guide's actions and stores them in the QRM. The QRM is characterized
by a \emph{semi-structured query} (SSQ), a schematic representation of
the information contained in the query, namely the inputs and output
contexts---the \emph{who, what, when,} and \emph{where}
categories---that characterize the question together with an
appropriately generalized representation of the guide's browser
interactions.  The \emph{query response evaluator} (QRE) evaluates a
query by binding specific arguments to its input parameters and
replaying the generalized web interactions the guide used to answer
the original question.

To answer a user's question, we must identify likely contextual
categories associated with the terms in the question to form an SSQ,
find those QRMs having contextually compatible SSQs, then use the QRE to
generate a response.

In what follows, we detail the elements of the system.  We first
discuss the processing of natural language queries to generate SSQs.
Next, we discuss the recording of guide behavior with the QRR.  We
then discuss how the categorical heterarchies associated with
ontological realms are constructed in a semi-automated fashion.  We
describe the method we use to assess how well a stored QRM matches a
user SSQ.  We then discuss the QRM evaluation process and discuss the
system's behavior on example queries. We conclude with a discussion of
future work.