### 2010-03-04 paper/semsearch10/content/introduction.tex

\section{Introduction}
% Christan
When traveling though a jungle to a destination, it is easy to get
lost.  The first person to go somewhere may make a number of mistakes
in trying to find the best path to their destination.  However, those
who come later will find it easy to reach that destination if a
well-marked trail without the has been created. Olsen and Malizia
describe this idea as \emph{exploiting trails} \cite{5379671}.  Rather
than treating each user's search as a unique entity and working from
first principles, one can exploit the fact that a similar search may
have been performed already.  In one study, almost 40 percent of all
queries were found to be repetitions of previous
queries.\cite{1277770} This points to way to optimize the search
process.  We have built a prototype system for answering questions
using deep-web sources that employs previously answered questions to
obtain new answers.  Using our analogy, the Morpheus question
answering system follows path-finders to their destinations and not
only marks the trail, but provides a taxi service to take followers to
similar destinations.