### 2010-03-04 paper/semsearch10/content/abstract.tex
latest version all of jnw & pete 1st changes made

\begin{abstract}
- It is difficult to answer questions that change over time
- However, if the method of answering the question can be realized, the question can be easily answered
- We are building a system that supports a learning how to answer particular questions
- We can answer questions that are semantically similar to our store of questions
- We do this by building a semantic library ...
\end{abstract}