### 2010-03-04 trunk/paper/semsearch10/content/introduction.tex
latest version all of jnw & pete 1st changes made

\section{Introduction}
% Christan
When traveling though a jungle to a destination, it is easy to get lost.  The first person to go somewhere may make a number of mistakes in trying to find the best path to their destination.  However, those who come later will find it easy to reach that destination if a well-marked trail has been created. Olsen and Malizia describe this idea as \emph{exploiting trails} \cite{5379671}.  Rather than treating a user's search as a unique entity and working without an example context, one can exploit the fact that a similar search may have been performed already.  In one study, almost 40 percent of all queries were found to be repetitions of previous queries\cite{1277770}. This indicates the reuse of prior searches is a way to optimize the search process.  We have built a prototype system for answering questions using deep-web sources that employs previously answered questions to obtain new answers.  Using our analogy, the Morpheus question answering system follows path-finders to their destinations and not only marks the trail, but also provides a taxi service to take followers to similar destinations.