### 2010-03-05 paper/semsearch10/content/nlp.tex
general editing

% Joir-dan stuff

When we have completed our Wikipedia corpora from all the blankets in the category hierarchy, our system needs to make sense of the data for execution. The learning process involves the calculation of two n-gram frequency distributions: an n-gram/vocabulary distribution and an n-gram/category distribution. The first value is the count of the individual n-grams in the combined vocabulary of all corpora. The second value is the count of the individual n-grams within a category's blanket. The categories corpora reflect one long context per category, instead of breaking up sentences into different contexts. The idea behind this is that Wikipedia pages are already grouped into categories, that we can assume are related to the topics of each sentence within the corpus. We store distributions, rather than priors, primarily because if at anytime a blanket vocabulary seems inadequate, we can append terms from other sources to both distribution types and recalculate our values. We can simply calculate probabilities and store these separately for the execution of our natural language query engine.

Our goal is to choose that category for use in the SSQ structure. This process requires the tagging and parsing of our natural language query, the extraction of key parts of speech, such as adjectives and nouns, and then using n-gram priors to calculate the appropriate category associated with each term.

Picking out these terms requires knowledge of the structure of the queThe ideal situation would be that we could recognize question patterns, and decide where key information is within the question based upon a given pattern.stion being passed to the system.  Currently, we only work with questions in an answer form, which means there is little to no transformation of the structure, to go from a question to an answer.

Consider the example question: "A 1997 Toyota Corolla needs what size tires?" The answer is: "A 1997 Toyota Corolla needs 17 inch tires."