### 2010-03-05 trunk/paper/semsearch10/content/qrr.tex

\section{Query Resolution Recorder (QRR)}
\label{sec:qrr}
% Chris
Answering a question using the deep web requires one to navigate
through a sequence of one or more web pages. Many of the accesses
involve clicks through web forms or links to resulting pages.  We have
developed a model that represents the various actions a user may carry
out during this process.

Web interactions can be broken down into three types:
\begin{enumerate}
\item The user fills in and submits a web form.
\item The user clicks a link.
\item The user highlights text that is needed in the query or forms
  part of the answer.
\end{enumerate}

When a users submit a web form they are prompted for the classes of
each of the form inputs. Here, class refers to categorization of data,
e.g. name, date, etc. These classes allow the QRR to map the user's
input SSQ tokens (which are also categorized) to their usage in a
given form. For a clicked link, the QRR simply records the target url
and nothing more. Finally, during the answer retrieval process when a
user discovers all or part of the desired information on a web page,
she highlights the text in question. This highlight allows the QRR to
record the source of the query's answer. All information gathered
during the recording process is saved in our data store.

Any query answering process consists of a sequence of these transactions.
Additionally, data collected for inputs and outputs may be found
during any stage of the complete query process.  An output
collected during one stage of an extraction may be the input to a
subsequent stage in the extraction.  The QRR builds the completed user
collection process into a QRM.

Conceptually, a QRM is the data structure that models the question
answering process.  Each QRM contains an ontilogical realm, an SSQ,
and information to support the query answering process.  The realm may
be one or more descriptive categories and is used for classification
of the QRMs.  The SSQ describes the set of possible input and output
categories from the associated natural language query.  The possible
set of phrase inputs and outputs are grouped into contextual bins
labeled \emph{who}, \emph{what}, \emph{when} and \emph{where}.  The
ontilogical realm and the SSQ parameters act as a signature for the
QRM object, analogous to a function signature.  The ontological
representation of the SSQ is described in section \ref{sec:ssq}.

Additionally, the QRM contains representations for every page visited
and the order of visitation.  For each page, we store the list of
inputs to the URL such as form inputs and the possible outputs
(highlighted portions of the webpage).  There is a mapping between the
SSQ inputs and outputs used on any page to the specific page input.
Finally, there is a structure with mappings to the user inputs or
outputs that are displayed to the user as answers.

The following is a mathematical description of the information
collected for user web interactions.

A QRM is a 4-tuple $Z = \left< \Sigma, P, M, A
\right>$ with the following structure:

\begin{enumerate}
%%%%%%%%%%%%%%% Separate the first 2 points as descriptions of the SSQ. --------------

\item $\Sigma = ( R, \Upsilon, \Omega )$ is the SSQ associated with the QRM.

  $R$ is the ontological realm associated with the QRM.

  $\Upsilon$ is the sequence of input categories drawn from $R$
  $\left< \Upsilon_{1}, \Upsilon_{2}, \Upsilon_{3}, \Upsilon_{4}
  \right>$ where $\Upsilon_{1}$ represents the \emph{who} context of
  the associated query, $\Upsilon_{2}$ represents the \emph{what}
  context, $\Upsilon_{3}$ represents the \emph{when} context, and
  $\Upsilon_{4}$ represents the \emph{where} context.

  $\Omega$ is the sequence of output categories drawn from $R$
  $\left<\Omega_{1}, \Omega_{2}, \Omega_{3}, \Omega_{4}\right>$ where
  $\Omega_{1}$ represents the \emph{who} context of the associated
  query, $\Omega_{2}$ represents the \emph{what} context, $\Omega_{3}$
  represents the \emph{when} context, and $\Omega_{4}$ represents the
  \emph{where} context.

\item $P$ is an ordered list of web pages $P = \left<P_1,P_2,...,
  P_n\right>$ where $P_h =
  \left<U_h,\left<I_{h_1},I_{h_2},...,I_{h_u}\right>,\left<O_{h_1},O_{h_2},...O_{h_v}\right>\right>$
  for $1 \leq h \leq n$, $u = \left| I_h \right|$, $v = \left| O_h
  \right|$. $P_h$ is a triple comprised of a URL $U_h$, a sequence of
  input arguments, and a sequence of output results.  \\ Let $I_{\$} =
  \bigcup_{j=1}^{4} I_{\downarrow_j}$ be the set of the selector
  expressions for the inputs of a query and $O_{\$} =
  \bigcup_{j=1}^{4} O_{\downarrow_j}$ be the set of selectors
  expressions for the outputs of a query.  \\ Let $K$ be the set of
  all string constants.
\begin{itemize}
\item URI $U_1 \in K$ is a string and \\ for $1 < g \leq n$, URI $U_g
  \in K \cup \left( \bigcup^{g-1}_{p=1} \left(
  \bigcup^{\left|O_p\right|}_{h=1}O_{p_h} \right)\right)$ where $1
  \leq h \leq \left| O_p \right| $,

\item for $1 < h < \left| I_1 \right|$, input $I_{1_h} \in K \cup
  I_{\$} \cup O_{\$}$ and \\ for $1 < g <= n$, then for $1 < h <
  \left| I_g \right|$, input $I_{g_h} \in K \cup I_{\$} \cup O_{\$}
  \cup \left( \bigcup^{g-1}_{p=1} \left(
  \bigcup^{\left|O_p\right|}_{h=1}O_{p_h} \right)\right)$.
\end{itemize}

\item $M$ is a map between $\Upsilon$, $K$, $\Omega$ to page list inputs $I_h$.

\item $A$ is the sequence of outputs from $\Upsilon$, $K$, and
  $\Omega$ representing the answer.

\end{enumerate}

%-------------------------------------------------------------------------------

%\subsection{Generalization}
% Christan

The response pages created by querying web forms are dynamically
generated with a template structure. This template structure contains
XPath [!!!!!need reference] selectors to access particular web page
elements.  One could employ XPath selectors that exactly match the
selection paths to the web page elements accessed by the guide during
query recording, however, using such specific paths fails to provide
robust performance because web sites are highly susceptible to
structural changes \cite{TanZMG07}.  To overcome this difficulty, we
generalize these XPaths using the \emph{GenPath algorithm} developed by
Badica et al. \cite{Badica06}.

Generalizing the extraction paths helps overcome small structural
changes. However, If a website changes to the point where a QRM is no
longer useful, the QRM is not considered fresh. User ratings of QRM
freshness can be employed in the SSQ matching process detailed in
Section \ref{sec:oqom}.

In many cases, the same information may be found by a number of guides
from the same deep web source, but by a different navigation path.
Therefore, we also employ a method of generalizing extraction paths
across page interactions. By doing this, one can remove unnecessary
page interactions and finds the most effective navigation path
necessary to produce a result.
%-------------------------------------------------------------------------------