### 2010-03-05 trunk/paper/semsearch10/paper.tex

\documentclass{www2010-submission}
\usepackage{listings}
\begin{document}

\title{Method Reuse for a Deep Web Question Answering System} 
\numberofauthors{1}
\author{
	 \alignauthor Christan Grant, Clint P. George, Joseph N. Wilson, Peter J. Dobbins, Jeff T. Depree, Christopher A. Shields and Joir-dan Gumbs\\
	 \affaddr{Dept. of Computer Science, University of Florida} \\ \affaddr
{ Gainesville, Florida, USA} \\
\email{\{cgrant, cgeorge, jnw, pjd, jepree, cshields, jgumbs\} @cise.ufl.edu}
}
\date{April 26,2010}

\maketitle

\input{content/abstract.tex}

\category{H.4.m}{Information Systems Applications}{Miscellaneous}

\keywords{deep web, ontology, question answering, re-finding}

\input{content/introduction.tex}
\input{content/example.tex}
\input{content/nlp.tex}
\input{content/qrr.tex}
\input{content/ontological_realms.tex}
\input{content/ssq_ranking.tex} % Joir-dan adds finding category stuff here
\input{content/query_reexecution.tex}
\input{content/results_futurework_conclusion.tex}

\bibliographystyle{abbrv}
\bibliography{citations}

\end{document}