### 2010-06-25 paper/IIWAS2010/content/related_work.tex
changes to introduction-ontology generation, category divergence, and results

\section{Related Work}
\label{sec:relatedwork}

\subsection{Question Ansering Systems}

Question Answering (QA) is an old field and there has been a recent influx of QA work. The earliest question answering systems include BASEBALL \cite{Green1961} and Lunar \cite{woods1973} were both \emph{closed domain} and \emph{closed corpus}.  Closed domain QA systems are those in which the possible meanings of the questions given are finite. We define closed corpus sytems as those in which the sources used to answer questions are finite.  QA systems that treat the web as a document collection is considered an open corpus system.  The morpheus system is considered an open corpus system.  It goes beyond current state of the art QA systems by examining deep web sources.

Several community QA systems have been developed, one of the most popular being Yahoo! Answers \cite{yahooanswers2008}.  These sites enlist a community of users to answer question posed by other users.  This idea has evolved to invent socially integrated QA systems have emerged such as Aardvark \cite{vark2010}. In addition to having questions answered by users, aardvark is able to contact potential answers through several mediums beyond its webpage (e.g. instant message, mobile phone).  This allows for a possibly faster answers to questions.  The morpheus system relies on both user interface and the property of similar question being answered by a community of users to give a fast reliable answer.

\subsection{Ontology generators} 
\label{sec:ontology_generators}

An ontology formally models real world \textit{concepts} and their
relationships. A concept or class in an ontology gives us an abstract and
simplified view of the world\cite{Gruber1993} in a machine-readable and
language-independent format. The relationships and attributes of an
ontological concept are defined using \textit{properties} and \textit{property
restrictions}. An ontology definition contains all of
these: classes, properties, and restrictions. A class in an ontology can have
multiple \textit{super classes} and \textit{sub-classes}. Finally, a
\textit{knowledge base} represents an ontology together with 
instances of its classes.

The DBpedia\footnote{http://dbpedia.org} project is based on extracting 
semantic information from the Wikipedia and making it available on the Web. 
Wikipedia semantics includes info-box templates, categorization information, 
images, Geo-coordinates, links to external Web pages, disambiguation pages, 
and redirects between pages in Wiki markup form \cite{Auer07dbpedia:a, Bizer2009}. 
DBpedia represents the Wikipedia categories using skos:concepts and category 
relations using skos:broader \cite{Auer07dbpedia:a, Bizer2009}. In fact, 
DBpedia does not define any new relations between the Wikipedia categories. 
The Wikipedia categories are extracted directly from Wikipedia pages, and there 
is no quality check on the resultant ontology. Similarly, additional ontological 
relations in DBpedia are generated from the info-box attributes and their values.

YAGO is a semi-automatically constructed ontology from Wikipedia and
WordNet\cite{Suchanek2009phd}. The YAGO ontology uses an automated process to
extract information from Wikipedia pages, info-boxes, and categories and to
combine this information with the WordNet\footnote{http://wordnet.princeton.edu}
synsets heterarchy. Since Wikipedia contains more individuals in the form of
Wikipedia pages than in the man-made WordNet ontology, the Wikipedia page titles
are used as the YAGO ontology individuals. YAGO concentrates mainly on the
fields of people, cities, events, and movies\cite{Suchanek2009phd}.

In addition, YAGO uses Wikipedia categories as its ontology classes. The
ontology's heterarchy is built using the \textit{hypernym} and \textit{hyponym}
relations of the WordNet synsets. YAGO uses only the nouns from WordNet and
ignores the WordNet verbs and adjectives. The connection between a WordNet
synset and a Wikipedia category is achieved by parsing the category names and
matching the parsed category components with the WordNet
synsets\cite{Suchanek2009phd}. Those Wikipedia categories having no WordNet
match are ignored in the YAGO ontology.

However, YAGO type extraction is mainly based on the assumption that
each Wikipedia page has at least one tagged category, and the assigned Wikipedia
categories are relevant to that Wikipedia page. The Wikipedia page category
assignments are subjectively assigned by human editors and some Wiki pages are
unassigned i.e., they have no category. In addition, we cannot completely rely
on the relevance of these assigned Wikipedia page categories.

\subsection{NLP Interfaces}