### 2010-06-29 paper/IIWAS2010/content/system_arch.tex

\section{System Architecure}
\label{sec:systemarch}

This section presents ontology and copora, query processing, ranking queries, and query executing.

\subsection{Using Ontology and Corpora} 
\label{sec:ontology_corpora}

Morpheus uses an ontology that contains categories of a particular realm of interest. Every leaf node in the ontology is associated with a corpus of words belonging to a category.  For example, we have constructed an automotive ontology containing categories relevant to the automotive realm. This ontology provides a structure for reference in the following sections.

Morpheus uses the DBpedia categories, Wikipedia pages and the WordNet synset heterarchy to construct ontologies. This approch is motivated by YAGO \cite{Suchanek2009phd}. First a realm is mapped to a DBpedia category \cite{Bizer2009}. 
Then a \emph{Markov Blanket} \cite{PRIS} is created covering all neighboring categories and using the DBpedia ontology properties \emph{broader} and \emph{narrower}. From here, the system recursively includes all categories up to the root.

Once we have the categories belonging to a particular realm, we associate them with WordNet synsets using an approach similar to YAGO\cite{Suchanek2009phd}. We determine a \emph{head noun} in the category name and search for it in the WordNet synsets. DBpedia categories having no WordNet synset match are excluded from our ontology.

To build a corpus for each of the leaf nodes in the ontology, 
we extract \emph{terms} from the Wikipedia pages associated with the 
the DBpedia categories. From this term corpus, 
we can find the likelihood of a term belonging to a particular category. This assists in categorizing terms in a \emph{path-follower} query.
The likelihood is determined by the probability of a category given a term 
using \textit{Bayes Rule} (Eq. \ref{eq:bayesrule}), since we can 
easily obtain the term-category and term-corpus probabilities as 
relative frequencies.

\begin{equation}
\label{eq:bayesrule}
P (category | term) = \frac{P(term | category) P(category)}{P(term)}
\end{equation}

In addition, we evaluate Latent Dirichlet Allocation (LDA) to identify latent topics of the documents in a corpus \cite{Blei2003latentdirichlet}.  LDA is Bayesian model represents a document in the corpus by distributions over topics, and a topic itself is a distribution over all  terms in the corpus.  For example, for the Wikipedia pages, the latent topics reflect the thematic structure of the pages. Thus, LDA can discover relevant topic proportions in a document using posterior inference \cite{Blei2003latentdirichlet}. Additionally, given a text document, using LDA we could tag related documents by matching the similarity over the estimated topic proportions, which is useful in the ontology and corpora building.  Due to its fully generative semantics, even at the level of documents, LDA could address the drawbacks (such as dimensionality and failure in finding the discriminative set of words for a document) of frequency based approaches such as TF-IDF, LSI, and pLSI.

\subsection{ Recording}
\label{sec:query_processing}

The \emph{Query Resolution Recorder} (QRR) is an instrumented web browser
used to track the search pathways to an answer for a \textit{path-finders} query. It
also helps guide to identify categories (optional) for the query terms. 
Morpheus has two types of categories: \emph{input categories} and \emph{output categories}.
Input categories annotates the query terms into a class or category, which are 
useful in finding similar queries. On the other hand, output categories give hints
to the expected answer of a user query. Morpheus captures all
this query information in the representation of an SSQ. For example, suppose, a guide 
tries to answer the query ``A 1997 Toyota Camry V6 needs what size tires?'', using
Morpheus. With the help of the QRR, guide may choose relevant query terms and assign
\emph{input categories} to them: Toyota : manufacturer, Camry V6 : 
model, tire : automotive parts, and size as the \emph{output category}. 
In addition, guide can log the search pathways to an answer (e.g. P215/60R16)
to this query and select the query \emph{realm} as automotive.

The \emph{Query Resolution Method} (QRM) is a data structure that models the
question answering process. A QRM is usually constructed by a \emph{guide} with
the help of the QRR. Each QRM contains an \emph{ontological realm}, an SSQ, and
information to support the query answering process. QRMs are associated with an
ontology that has a particular realm i.e. an ontological realm. In addition, 
the QRM contains information required to visit each page needed to 
answer the query as well as the order of visitation. For each
dyanamic page, it stores the list of inputs to the URL such as form inputs and
the possible referrence outputs (highlighted portions of the web page).

% NLP arch. 
% Joir-dan

Rather than working blindly, forming all possible n-grams from the query 
and running probabilities, it seemed to be a better approach to identify 
important noun phrases within the query, and work with those. Using our 
query parser, we proceed to tokenize, parse, and chunk our candidate query, 
extracting verb groups, answer classes, and descriptive info from each query. 
In the example query "What is the tire size of a 1997 Toyota Camry?," 
the verb group is the verb "is," the answer class is "tire size," and descriptive 
info is "of a 1997 Toyota Camry." The next step is to create n-grams from these chunks. 
Using these n-grams, we will query our database for frequencies associated with 
each n-gram: overall frequency and categorical frequency, alongside category 
term count and overall term count, to calculate categorical probabilities 
and establish a realm for the candidate query, choosing realm with highest probability.

\subsection{Ranking} 
\label{sec:qrm_ranking}

To answer a user's query, a \textit{candidate
SSQ}, Morpheus finds similar SSQs that belong to QRMs in the Morpheus data store (a \textit{qualified SSQ}). We need a similarity measure to match the candidate SSQ
with a qualified SSQ. For the SSQ similarity, we
consider the SSQ components such as query-realm, input
terms and output terms, and their assigned categories or classes [To be changed]. The,
\textit{category divergence} of two categories characteristics their 
similarity based on an ontology class heterarchy. 
It is motivated by the concept of multiple
dispatch in CLOS and Dylan programming for generic function \emph{type} matches.

We consider the category match as a type match and we use 
\emph{category divergence} to calculate the relevance between 
the candiate SSQ and a qualified SSQ. Each qualified SSQ will 
have input terms, output terms, and associated classes, and one realm (from QRM). 
For the candidate SSQ, the relevant categories for terms 
are determined from the NLP engine and corpora. The calculation of a realm 
for a candidate query is done using the n-grams found within said query 
and the probabilities (if any) found using frequencies within the database.
We match QRMs that belong to the same \emph{realm} of the candidate SSQ. 
The relevance of a qualified SSQ to the candidate SSQ is 
determined by aggregating the divergence measure of input-term-categories 
associated with them. In addition, we order QRMs in the data store 
by descreasing relevance. The order provides a ranking for the results 
to the user. The following subsection describes \emph{category divergence} 
in detail.

\subsubsection{Catgeory Divergence}
\label{sec:ctd}

We employ a quasimetric, \textit{category
divergence (cd)},
between a source category and a target category using the \textit{topological
structure} of the categories in an ontology. We write $S \prec T$ for the
reflexive
transitive closure of the supercategory relation. Let $d(P,Q)$ represent the hop
distance in the directed ontology inheritance graph from $P$ to $Q$. The
divergence between a source and target category ranges from zero (for identical
categories) to one (for type incompatible categories). Let $S$ be the source
category, $T$ be
the target category, and $C$ be a least common ancestor category of $S$ and
$T$ i.e., one that minimizes $d(S,C) + d(T,C)$. The category divergence between $S$ and $T$ is defined given by:

\begin{equation}
cd(S, T) = \begin{cases}
0 & S.{Uri} \equiv T.{Uri}\\
d(S, T)/(3h) & S \prec T\\
1 & T \prec S\\
(d(S,root) + d(S,C) \\ \ \ \ \ + d(T,C))/(3h) & otherwise
\end{cases}
\end{equation}

\noindent where $h$ is the height of the ontology tree.

Note, if $S \prec T$ and $S \not\prec Q$ then $cd(S,T) <
cd(S,Q)$, that is, the divergence of a source category to a target
ancestor category is smaller than the divergence of a source category to any
category that is not an ancestor. This is an important property in
determining the compatibility of categories for answering queries.  If a
SSQ answers queries concerning an ancestor category, it is more relevant
that a SSQ that answers queries from any non-ancestral category.

% Algorithm 1: figure  
\begin{figure}[t]
\centering
\includegraphics[width=85mm]{img/automotive_ontology.eps}
\caption{An abbreviated man-made automotive ontology - arrows represent inheritance relationships between the ontological classes.}
\label{fig:automotive_ontology}
\end{figure}

Suppose we want to find the category divergence between \textit{Sedans}
and \textit{Trucks} (Figure \ref{fig:automotive_ontology}). 
\textit{Sedans} is the subcategory of \textit{Cars}, and \textit{Trucks}
is the subcategory of \textit{Automobiles}. Therefore, the least common ancestor ($C$)
for these twoThis section presents ontology and copora, Query processing, ranking queries, and the query resolution mechanism. categories is \textit{Automobiles}. In addition, the tree height \textit{h = 4}.
So, the normalized divergence $cd$\textit{(Sedans, Trucks)} is \textit{7/12}
that is calculated from \textit{d(Sedans, Root) = 4}, \textit{d(Sedans, Automobiles) = 2}, and \textit{d(Trucks, Automobiles) = 1}.

\subsection{Executing}

QRM contains the necessary  
information required to re-run this procedure.Once we have 
relevant QRMs through QRM ranking for a given user query, we can  
get answers by re-running the pathways with help of 
the \emph{Query Resolution Engine} (QRE) of Morpheus.