### 2010-07-01 paper/IIWAS2010/content/introduction.tex

\section{Introduction}

% Outline 
% ---------------------

% 1. Motivation 
% 2. What are we proposing?  
% 3. Challenges
% 		3.1 User query processing 
% 		3.1 Need of an ontology 
% 4. Realm based question answering 
% 5. About the paper structure

When traveling though a jungle to a destination, it is easy to get lost.  The first person to journey somewhere may make a number of mistakes when trying to find the best path to their destination. Those who come later find it easier to reach the destination if a well-marked trail has been created. Olsen and Malizia describe this idea as \emph{exploiting trails} \cite{5379671}.  Rather than treating a user's discovery experience as a unique entity, one can exploit the fact that a similar search may have already been performed.  In one study, almost 40 percent of all queries were repetitions of previous queries\cite{1277770}. Thus, reuse of prior searches is one way to optimize the search
process.  Morpheus is a question answering system motivated by reuse of prior web search pathways to yield an answer to a user query. Morpheus follows path-finders to their destinations and not only marks the trail, but also provides a taxi service to take followers to similar
destinations.

There are two distinct Morpheus user roles. A
\textit{path-finder} enters queries in the Morpheus web interface and
searches for an answer to the query using an instrumented web browser. 
This web tracking tool stores the query
and necessary information to revisit the pathways to the page where the path-finder 
found the answer. A \textit{path-follower} uses the Morpheus system much like a regular search engine with a natural language interface. The path-follower enters a question in a text box and receives a guided path to the answer. The system exploits previously found paths to provide an answer.

Morpheus represents user-query details in a semi-structured query (SSQ). It assumes the terms belong to classes of a \textit{consistent} realm-based ontology, that is, one having a singly rooted heterarchy whose subclass/superclass relations have meaningful semantic interpretations. When a path-follower enters a query, Morpheus ranks SSQs in the store based on class similarity. Suppose a regular user asks -\textit{ a 1997 Toyota Camry V6 needs what size tires?} In this query the classes associated with terms, e.g. \emph{Manufacturer} with \emph{Toyota}, helps us identify similar queries.

This paper discusses related question answering and ontology generation systems in section \ref{sec:relatedwork}. Section \ref{sec:systemarch} explains the Morpheus system and its implementation. In Section \ref{sec:results} we describe the current results of our approach.  Finally, we conclude with future goals for the system.